---
title: "testf-vignette"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
    math: katex
    df_print: paged
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{testf-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(F2020Projt3LP)
```

# Purpose

The intent of this vignette is for me to experiment with making a vignette by creating a short description of my testf() function within the package

# testf

## Why does it exist

I wrote this function as a way to check that I had set up my package properly. I wanted to make sure that everything was working, especially with GitHub, before putting the important functions into the package

## What does it do

It is a very simple function that will take a vector input, square it, return the output, and then also return a statement saying what the purpose of the function is. This is not important for the graded/assigned part of Project 3.

## Example

Here is an example of the function working

```{r}
x <- 1:10
testf(x)
```

As you can see from the above output, the function does what I described.
